Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic

### 2021-09-15 psy_peds.Rmd
correcting stats based on age at the time of visits

all_codes <- obs_raw %>% 
  full_join( clinical_raw,  by = c("patient_num", "cohort", "days_since_admission")) %>%
  full_join( demo_raw, by = c("patient_num", "cohort")) %>% 
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others")) %>% 
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 in_hospital == 1,
                 age < 18 & age > 10,
                 date >= history_start_date & date <= history_end_date) %>%
  dplyr::select( patient_num, cohort, sex, age, concept_type, concept_code, calendar_date,
                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
                 period, disorder_group, description, icd_code_category )

@@

```{r}
#estimate the number of patients per week by age group
input_data[, c("patient_num", "cohort", "age","weeks")] %>%
  unique() %>%
  mutate( age_group = ifelse( age < 15, "early adolescence (11-14 years)", "middle adolescence (15-17 years)")) %>%
  group_by(weeks, age_group) %>%
  tally() %>%
  mutate(n = ifelse( n > obfuscation, n, obfuscation * 0.5))  %>%
    ggplot(aes(x = as.Date(weeks), y = n, fill = age_group)) +
    geom_bar(stat="identity", position = "dodge")+
    labs(y = "Counts",
         x = "Week",
         title = "Weekly patient counts with mental health related ICD codes aggregated by age group",
         subtitle = "Per age group") + 
    theme(legend.position = "bottom") + 
  scale_fill_manual(values = cbPalette) +
    geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
ggsave("output/overall_weekly_count_age.png", width = 18, height = 9, units="cm")

@@

## Age
```{r}
age_stats <- input_data %>%
  dplyr::distinct(patient_num, cohort, period, age) %>%
  filter(age != 0) %>% 
  dplyr::group_by( period ) %>% 
  dplyr::summarise( mean = mean( age ), 
                    sd = sd( age ), 
                    count = n()) %>% 
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::ungroup() %>% 
  dplyr::bind_rows(
    dplyr::summarise(input_data, period="Total", mean=mean(age), sd=sd(age), count = n())
  )

>>>>

all_codes <- obs_raw %>% 
  full_join( clinical_raw,  by = c("patient_num", "cohort", "days_since_admission")) %>%
  full_join( demo_raw, by = c("patient_num", "cohort")) %>% 
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others")) %>% 
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 in_hospital == 1,
                 age_time_visit < 18 & age_time_visit > 10,
                 date >= history_start_date & date <= history_end_date) %>%
  dplyr::select( patient_num, cohort, sex, age, age_time_visit, concept_type, concept_code, calendar_date,
                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
                 period, disorder_group, description, icd_code_category )

@@

```{r}
#estimate the number of patients per week by age group
input_data[, c("patient_num", "cohort", "age_time_visit","weeks")] %>%
  unique() %>%
  mutate( age_group = ifelse( age_time_visit < 15, "early adolescence (11-14 years)", "middle adolescence (15-17 years)")) %>%
  group_by(weeks, age_group) %>%
  tally() %>%
  mutate(n = ifelse( n > obfuscation, n, obfuscation * 0.5))  %>%
    ggplot(aes(x = as.Date(weeks), y = n, fill = age_group)) +
    geom_bar(stat="identity", position = "dodge")+
    labs(y = "Counts",
         x = "Week",
         title = "Weekly patient counts with mental health related ICD codes aggregated by age group",
         subtitle = "Per age group") + 
    theme(legend.position = "bottom") + 
  scale_fill_manual(values = cbPalette) +
    geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
ggsave("output/overall_weekly_count_age.png", width = 18, height = 9, units="cm")

@@

## Age
```{r}
age_stats <- input_data %>%
  dplyr::group_by(patient_num, cohort, period) %>% 
  dplyr::filter(age_time_visit == max(age_time_visit)) %>%
  dplyr::ungroup() %>% 
  dplyr::distinct(patient_num, cohort, period, age_time_visit) %>% 
  dplyr::filter(age_time_visit != 0) %>% 
  dplyr::group_by( period ) %>% 
  dplyr::summarise( mean = mean( age_time_visit ), 
                    sd = sd( age_time_visit ), 
                    count = n()) %>% 
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::ungroup() %>% 
  dplyr::bind_rows(
    dplyr::summarise(input_data,
                     period="Total",
                     mean = mean(age_time_visit),
                     sd = sd(age_time_visit),
                     count = n())
  )
